### 2015-05-16 Draft/conclusion.tex
Updated Dr. Xie's Comments

\section{Conclusion}
\label{sec:conclusion}
\vspace*{-1ex}
Despite being highly desirable, formal temporal constraints are absent in most APIs.
In contrast, documentation of API methods contains detailed specifications of temporal constraints in natural language text.
Manually writing formal specifications based on natural language text in API documents is prohibitively time-consuming and error-prone.
To address this issue, we have proposed a novel approach called \tool\ to infer temporal constraints from natural language text of API documents.
We used \tool\ to infer temporal constraints from
the \paypalAPI, the \amazonAPI, and the 
commonly used package \CodeIn{java.io} in the JDK API.
Our evaluation results show that \tool\ effectively identifies sentences describing
temporal constraints with an average 79\% precision and 60\% recall,
from more than 4000 sentences in subject API documents.
Furthermore, \tool\ also achieves an accuracy of
70\% in inferring 77 formal temporal constraints from these temporal constraint sentences.